Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics

### 2023-03-30 03_marlin_paper.qmd
merging back to main prior to submission

---
title: "Revealing Benefits, Costs, and Tradeoffs of Spatial Management in Marine Systems Arrising from Coupled Economic and Ecological Dynamics"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Echelle Burns
    - Lennon Thomas
    - James Thorson
date: "Updated `r Sys.Date()`"
format:
  pdf:
    toc: true
    fig-pos: 'H'
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
bibliography: references.bib
params:
  run_name: ["v1.0"]
abstract: "Abstract: Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research. We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets. This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics."
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

@@

@@

# Data Availability Statement

All code and materials needed to reproduce this paper is available at (https://github.com/DanOvando/marlin-paper)\[https://github.com/DanOvando/marlin-paper\].

# References

@@

# Supporting Information

>>>>

---
format:
  pdf:
    toc: false
    fig-pos: 'H'
    mainfont: Times New Roman
    sansfont: Times New Roman
    fontsize: 12pt
  html: default
  docx: default
execute:
  echo: false
  message: false
  warning: false
  cache: false
csl: chicago-author-date-16th-edition.csl
bibliography: references.bib
params:
  run_name: ["v1.0"]
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
---

@@

## Title Page

Journal Name: Ecological Applications

Manuscript Type: Article

Manuscript Title: Simulating Benefits, Costs, and Tradeoffs of Management Policies in Marine Social-Ecological Systems

Author Names: Daniel Ovando^1,2^, Darcy Bradley^3,4,5^, Echelle Burns^3,4,5^,Lennon Thomas^3,4,5^, James Thorson^6^

Affiliations:
^1^ School of Aquatic and Fishery Sciences, University of Washington,1122 NE Boat St Box 355020,Seattle, WA, USA
^2^ Current Address: Inter-American Tropical Tuna Commission,8901 La Jolla Shores Drive, CA, USA
La Jolla, CA 92037, USA
^3^ Marine Science Institute, University of California, Santa Barbara, CA, USA
^4^ Bren School of Environmental Science & Management, University of California, Santa Barbara, CA, USA
^5^ Environmental Markets Lab, University of California, Santa Barbara, CA, USA
^6^ Habitat and Ecological Processes Research Program, Alaska Fisheries Science Center, NOAA, NMFS, NOAA, Seattle, WA, USA

Corresponding Author: Daniel Ovando <dovando@iattc.org>

Open Research Statement: All code and materials needed to reproduce this paper is available at [https://github.com/DanOvando/marlin-paper](https://github.com/DanOvando/marlin-paper).

Key Words: 
Bio-economic modeling, spatio-temporal modeling, marine protected areas, fisheries management, marine spatial planning, conservation planning

\newpage

## Abstract

Designing effective spatial management strategies is challenging because marine ecosystems are highly dynamic and opaque, and extractive entities such as fishing fleets respond endogenously to ecosystem changes in ways that depend upon ecological and policy context. We present a tool, `marlin`, that can be used to efficiently simulate the bio-economic dynamics of marine systems in support of both management and research. We demonstrate `marlin’s` capabilities by focusing on two case studies on the conservation and food production impacts of marine protected areas (MPAs): a coastal coral reef and a pelagic tuna fishery. In the coastal coral reef example, we explore how heterogeneity in species distributions and fleet preferences can affect distributional outcomes of MPAs. In the pelagic case study, we show how our model can be used to assess the climate resilience of different MPA design strategies, as well as the climate sensitivity of different fishing fleets. This paper demonstrates how intermediate complexity simulation of coupled bio-economic dynamics can help communities predict and potentially manage trade-offs between conservation, fisheries yields, and distributional outcomes of management policies affected by spatial bio-economic dynamics.

@@

Users can manually supply a vector of of natural mortality at age (*m*).
Or, they can supply one value of natural mortality which is then
converted into mortality at age through one of two means. Under the
default behavior, natural mortality at age given a target mean mortality
across all ages $m_s$ is calculated using a length-inverse mortality
function [@lorenzen2022].

$$
minv_{s_a} = (\frac{l_{s,a}}{l_{\infty,s}})^{-1}
$$ {#eq-minv}

$$
m_{s,a} = \frac{minv_{s,a}}{mean( minv_{s,a})}m_s
$$ {#eq-mata}

Alternatively, users can set mortality at age to be constant

$$
m_{s,a} = m_s
$$ {#eq-mata2}

@@

## Acknowledgements

This project was funded by the Waitt Foundation. We thank Dr. Alexa Fredston for helpful revisions of this manuscript.

## Author Contribitions

Methods and code written by DO and JT. Results produced by DO. Data collected by EB and LT. All authors contributed to the conceptualization and writing of the manuscript.

## Conflit of Interest Statement

The authors declare no conflicts of interest.

## References

@@
